---
title: "QPM II - Class Session 17"
output: pdf_document
date: "2024-10-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("sjPlot")) install.packages("sjPlot", dependencies = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
library(sjPlot)
```

1. The demo video does not remove records for which the feeling thermometers are not in the range between 0 and 100.
• How many records do not have appropriate feeling thermometer values?
• Remove those records and run the logistic regression again. Check if there is any difference in the summary.

```{r}
data <- read_csv("dataset-anes-2012-subset1.csv")

# Count number of records with invalid thermometer values
num_invalid_records <- data %>%
  filter(!(ft_dem >= 0 & ft_dem <= 100 &
           ft_rep >= 0 & ft_rep <= 100)) %>%
  nrow()

print(paste("Number of records with invalid feeling thermometer values:", num_invalid_records))

# Remove rows with invalid thermometer values
data_clean <- data %>%
  filter(ft_dem >= 0 & ft_dem <= 100,
         ft_rep >= 0 & ft_rep <= 100)

# Run logistic regression
logistic_model <- glm(vote_obama ~ ft_dem + ft_rep + income + sex + other + black + hispanic,
                      data = data_clean, family = "binomial")

# Print the summary of the model
summary(logistic_model)
```
2. Draw the logistic regression plot with 95% confidence intervals and the accompanying rug representa- tion. The independent variable in question is ft_dem. Hold all other variables at the mean value. The plot will look like this:
```{r}
# c(0, 100) equal interval
ft_dem_seq <- data.frame(ft_dem = seq(min(0, na.rm = TRUE), max(100, na.rm = TRUE), length.out = 100))

# mean of other predictor variables
mean_values <- data_clean %>%
  summarise(across(c(ft_rep, income, sex, other, black, hispanic), mean, na.rm = TRUE))

# bind(seq, mean_values)
predict_data <- ft_dem_seq %>%
  bind_cols(mean_values)

# Predicted probs & standard errors
predictions <- predict(logistic_model, newdata = predict_data, type = "response", se.fit = TRUE)

# Compute confidence intervals
predict_data <- predict_data %>%
  mutate(predicted_prob = predictions$fit,
         lower = predictions$fit - 1.96 * predictions$se.fit,
         upper = predictions$fit + 1.96 * predictions$se.fit)

ggplot(predict_data, aes(x = ft_dem, y = predicted_prob)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  labs(x = "Feeling Thermometer for Democrats (ft_dem)", y = "Predicted Probability") +
  theme_minimal()

# Can add the rug, but probably only works in base R.

```


3. Draw the marginal effect plots with 4 terms: income (the values on x-axis), ft_rep, race, and sex. You
may need to install the mfx package.
• What patterns do you see?
• Redraw the plot to limit the ft_rep to be [0, 25, 50, 75, 100], race to be white and other, and sex
to be female only.

```{r}
logistic_model2 <- glm(vote_obama ~ ft_rep + income + sex + race,
                      data = data_clean, family = "binomial")

plot_model(logistic_model2, type = "pred", terms = c("income", "ft_rep", "race", "sex"))

# Now with filtered data
filtered_data <- data_clean %>%
  filter(ft_rep %in% c(0, 25, 50, 75, 100),
         race %in% c(1, 4),
         sex == 1)

filtered_logistic_model2 <- glm(vote_obama ~ ft_rep + income + sex + race,
                      data = filtered_data, family = "binomial")

plot_model(filtered_logistic_model2, type = "pred", terms = c("income", "ft_rep", "race", "sex"))


```
