---
title: "ClassSession14"
output: html_document
date: "2024-10-27"
---

## Assuming the null hypothesis is true, simulate n = 40 with sigma = 2. Calculate p-value.

```{r}
set.seed(12345)
n = 40
mu = 0
sigma = 2
null_mu = 0

data = rnorm(n, mean = mu, sd = sigma)
print(1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n)))
```

## Repeat 400 times

```{r}
N = 400
n = 40
mu = 0
sigma = 2
null_mu = 0

ps = rep(0, N)
for (i in 1:N){
  data = rnorm(n, mean = mu, sd = sigma)
  ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
}
plot(density(ps))
```

## Assume that the null hypothesis is False. What percent of the time do you reject? What is that called?
### Power! Reject 94% of the time. 
```{r}

N = 400
n = 40
mu = 1
sigma = 2
null_mu = 0

ps = rep(0, N)
for (i in 1:N){
  data = rnorm(n, mean = mu, sd = sigma)
  ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
}

plot(density(ps))

print(sum(ps<=0.05)/N)
```

## Range of values for the true μ. At each value, calculate the probability of rejecting. Plot that. What’s that called?
### Power Curve!
```{r}
num_mus=20
mus = seq(from = 0, to = 2, length.out = num_mus)
powers = rep(0, num_mus)

for(k in 1:num_mus){
  N = 400
  n = 40
  mu = mus[k]
  sigma = 2
  null_mu = 0
  ps = rep(0, N)
  
  for (i in 1:N){
    data = rnorm(n, mean = mu, sd = sigma)
    ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
  }
  
  powers[k] = sum(ps<=0.05)/N
}
plot(1:num_mus, powers)
```

## now fix μ and alter the sample size.
```{r}

M = 20
mu = 1
ns = seq(from = 10, to = 200, length.out = M)
powers = rep(0, M)

for(k in 1:M){ 
  N = 400
  n = ns[k] 
  sigma = 2 
  null_mu = 0
  ps = rep(0, N)
  
  for (i in 1:N){
    data = rnorm(n, mean = mu, sd = sigma)
    ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
  }
  
  powers[k] = sum(ps<=0.05)/N
}
plot(1:M, powers)
```

## now alter σ2
```{r}
M = 20
mu = 1
sigmas = seq(from = 1, to = 10, length.out = M)
powers = rep(0, M)

for(k in 1:M){ 
  N = 400
  n = 40
  sigma = sigmas[k]
  null_mu = 0
  ps = rep(0, N)
  for (i in 1:N){
    data = rnorm(n, mean = mu, sd = sigma)
    ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
  }
  
  powers[k] = sum(ps<=0.05)/N
}
plot(1:M, powers)
```

## Drug Dosage Problem
```{r}
p_vec = c(0.1, 0.4, 0.5, 0.6, 0.7, 0.8)
power_vec = c(1, 0.979, 0.86, 0.584, 0.228, 0.032)

plot(p_vec, power_vec, type="l", xlab = "Probability, p", ylab = "Power")

```
