---
title: "QPM II - Class Session 25"
output: pdf_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
library(readr)
library(haven)
library(ggplot2)
library(dplyr)
library(glmnet)
```

(1) Using the provided dataset communities, we are going to fit OLS, LASSO, and ridge regression and evaluate each model’s performance. First, separate the data into a training set, a validation set, and a test set. Use 1500 observations for the training set. Set aside 247 for the validation set and the remaining 247 for the test set.
```{r}
load("communities.Rdata")
communities <- cbind(X)

set.seed(12345) 

# Shuffle rows
communities <- communities %>%
  sample_frac(1) 

# Split dataset
train_set <- communities %>% slice(1:1500)
validation_set <- communities %>% slice(1501:1747)
test_set <- communities %>% slice(1748:1994)

x_train <- train_set %>% select(-ViolentCrimesPerPop) %>% as.matrix()
y_train <- train_set$ViolentCrimesPerPop

x_val <- validation_set %>% select(-ViolentCrimesPerPop) %>% as.matrix()
y_val <- validation_set$ViolentCrimesPerPop

x_test <- test_set %>% select(-ViolentCrimesPerPop) %>% as.matrix()
y_test <- test_set$ViolentCrimesPerPop

```

(2) Now use the training set to fit several candidate models to predict the outcome variable ViolentCrimesPerPop. Specifically, fit OLS with a set of variables (PctUnemployed, PctLess9thGrade, PctPopUnderPov, PctWorkMom) which theoretically should best predict violent crime, and use the glmnet package to fit LASSO and Ridge with all variables except the outcome (which is the last column of the dataset).
```{r}

ols_predictors <- c("PctUnemployed", "PctLess9thGrade", "PctPopUnderPov", "PctWorkMom")

# Data
train_set_ols <- train_set %>% select(all_of(ols_predictors), ViolentCrimesPerPop)
val_set_ols <- validation_set %>% select(all_of(ols_predictors), ViolentCrimesPerPop)
test_set_ols <- test_set %>% select(all_of(ols_predictors), ViolentCrimesPerPop)

# OLS
ols_model <- lm(ViolentCrimesPerPop ~ ., data = train_set_ols)

# LASSO model (alpha = 1 for LASSO)
lasso_model <- cv.glmnet(x_train, y_train, alpha = 1, standardize = TRUE)

# Ridge model (alpha = 0 for Ridge)
ridge_model <- cv.glmnet(x_train, y_train, alpha = 0, standardize = TRUE)


```

(3) With the validation set, estimate prediction error using mean squared error for each of the candidate models.
(4) With the test set, assess the generalizability of the models. Can it predict the new data well?

```{r}

ols_predictions <- predict(ols_model, newdata = val_set_ols)
lasso_predictions <- predict(lasso_model, s = "lambda.min", newx = x_val)
ridge_predictions <- predict(ridge_model, s = "lambda.min", newx = x_val)

# Evaluate performance on validation
ols_rmse <- sqrt(mean((y_val - ols_predictions)^2))
lasso_rmse <- sqrt(mean((y_val - lasso_predictions)^2))
ridge_rmse <- sqrt(mean((y_val - ridge_predictions)^2))

# Print RMSE results
print(paste("OLS RMSE:", ols_rmse))
print(paste("LASSO RMSE:", lasso_rmse))
print(paste("Ridge RMSE:", ridge_rmse))

# test set
ols_predictions_tst <- predict(ols_model, newdata = test_set_ols)
lasso_predictions_tst <- predict(lasso_model, s = "lambda.min", newx = x_test)
ridge_predictions_tst <- predict(ridge_model, s = "lambda.min", newx = x_test)

# Test set
ols_rmse <- sqrt(mean((y_test - ols_predictions)^2))
lasso_rmse <- sqrt(mean((y_test - lasso_predictions)^2))
ridge_rmse <- sqrt(mean((y_test - ridge_predictions)^2))

# Print RMSE results
print(paste("OLS RMSE (test):", ols_rmse))
print(paste("LASSO RMSE (test):", lasso_rmse))
print(paste("Ridge RMSE (test):", ridge_rmse))

```
```{r}

# Prep data
train__and_val_set <- communities %>% slice(1:1747)
x_data <- train__and_val_set %>% select(-ViolentCrimesPerPop) %>% as.matrix()
y_data <- train__and_val_set$ViolentCrimesPerPop

# split into 10 folds
folds <- sample(1:10, size = nrow(x_data), replace = TRUE)

# Initialize a vector to store MSE for each fold
rmse_values <- numeric(10)

for (i in 1:10) {
  
  # Data split
  test_indices <- which(folds == i)
  x_train <- x_data[-test_indices, ]
  y_train <- y_data[-test_indices]
  x_test <- x_data[test_indices, ]
  y_test <- y_data[test_indices]
  
  # Fit and predict
  lasso_model <- cv.glmnet(x_train, y_train, alpha = 1, standardize = TRUE)
  lasso_predictions <- predict(lasso_model, newx = x_test, s = "lambda.min")
  
  # Calculate MSE
  rmse_values[i] <- sqrt(mean((y_test - lasso_predictions)^2))
}

# MSE average
average_rmse <- mean(rmse_values)
print(paste("Average RMSE from 10-fold cross-validation:", round(average_rmse, 4)))


```