---
title: "QPM II - Class Session 15"
output: pdf_document
date: "2024-10-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We are going to think about hypothesis testing again. In particular, we are going to think about testing:

H0 : μ = 0
H1 : μ = 20

This is somewhat contrived, but the example will help you think through some of the differences between all of the various approaches.

Use this code to generate the data:
set.seed(1234)
x<-rnorm(60, mean=3, sd=1)
x.bar<-mean(x)
sd<-sd(x)

We will assume that σ2 is fixed and known (so not a parameter).

1. Calculate the p-value for this test. What conclusion did you reach? Does that seem right?


```{r}
set.seed(1234)
x <- rnorm(60, mean=3, sd=1)
x.bar <- mean(x)
sd <- sd(x)

# Assuming known standard deviation from sample (σ)
n <- 60
mu_0 <- 0
sigma <- sd

pnorm(x.bar, mean = 0, sd = sd/sqrt(n), lower.tail=FALSE)
```

You ignored the alternative hypothesis. You decided that there are only two possible worlds. Anything against null ~ for the alternative. Proof by contradiction. ~ Logical Fallacy

2. Calculate the likelihood ratio test for this hypothesis. What conclusion do you reach? Does that seem right?
```{r}
# Likelihood under H0: mu = 0
L_H0 <- exp(sum(dnorm(x, mean = 0, sd = sigma, log = TRUE)))

# Likelihood under H1: mu = 20
L_H1 <- exp(sum(dnorm(x, mean = 20, sd = sigma, log = TRUE)))

# Calculate the Likelihood Ratio
lrt <- L_H0 / L_H1
lrt
```

exp(sum(log(probabilities))) due to vanishing gradient.

3. Calculate the Bayes factor for this hypothesis. What conclusion do you reach? Does that seem right?
```{r}

# Calculate Bayes Factor
BF <- L_H1 / L_H0

# Prior odds (2 to 1 against H1 means prior odds of H1 = 1/2)
prior_odds <- 1 / 2

# Posterior odds
posterior_odds <- BF * prior_odds

# Output Bayes Factor and Posterior Odds
list(BF = BF, Posterior_Odds = posterior_odds)

```
4. Now repeat (1) but use x<-rnorm(60, mean=1, sd=.1)

Is this evidence better or worse for the alternative hypothesis?
```{r}
# Step 4: Generate new data with mean = 1, sd = 0.1
set.seed(1234)
x <- rnorm(60, mean=1, sd=.1)
x.bar <- mean(x)
sd <- sd(x)

# Assuming known standard deviation from sample (σ)
n <- 60
mu_0 <- 0
sigma <- sd

pnorm(x.bar, mean = 0, sd = sd/sqrt(n), lower.tail=FALSE)
```