---
title: "QPM II - Class Session 26"
output: pdf_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
library(haven)
library(dplyr)
library(sjPlot)
library(ggplot2)
df <- read.csv("gender_data.csv")
```

Fit an OLS model using pol_participation as the outcome variable. Include a two-way interaction term of age and gender. Write down the model specification in latex. Plot the interaction effects. Are the marginal effects different between males and females? Run the summary() function on your model. Interpret the plot and the coefficients.

```{r}

# Fit the OLS model
model <- lm(pol_participation ~ age * gender + higher_edu + income_chng + polinterest +
              home_child + income + infection + self_esteem + pol_participation_past, 
            data = df)

# Display the summary of the model
summary(model)

# Plot interaction effects
plot_model(model, type = "int", terms = c("age", "gender")) +
  labs(title = "Interaction Effect of Age and Gender on Political Participation",
       x = "Age", y = "Political Participation") +
  theme_minimal()

```
```{r}

# Fit the model with three-way interaction
model_three_way <- lm(pol_participation ~ age * gender * polinterest + higher_edu + 
                      income_chng + home_child + income + infection + 
                      self_esteem + pol_participation_past, data = df)

# Summary of the model
summary(model_three_way)

```
Create a new column change_participation which is the difference between pol_participation
and pol_participation_past. Use change_participation as the outcome variable. Form your own
hypothesis. Fit an OLS model with predictors and interaction terms of your choice. Plot the marginal
effects and interpret the results.
```{r}

df$change_participation <- df$pol_participation - df$pol_participation_past

model_change <- lm(change_participation ~ polinterest * gender + self_esteem + 
                   age + higher_edu + income_chng + home_child + 
                   infection + income, data = df)

# Summary of the model
summary(model_change)

plot_model(model_change, type = "pred", terms = c("polinterest", "gender")) +
  labs(title = "Marginal Effects of Political Interest on Change in Participation",
       x = "Political Interest", y = "Change in Political Participation",
       color = "Gender") +
  theme_minimal()

```
```{r}
df$age_level <- ifelse(df$age < median(df$age, na.rm = TRUE), "young", "old")
df$age_level <- as.factor(df$age_level)

model_logit <- glm(gender ~ age_level * pol_participation + higher_edu + 
                   income_chng + home_child + infection + income + 
                   self_esteem + pol_participation_past, 
                   family = binomial(link = "logit"), data = df)

# Summary of the model
summary(model_logit)

plot_model(model_logit, type = "pred", terms = c("pol_participation", "age_level")) +
  labs(title = "Marginal Effects of Political Participation and Age Level on Gender",
       x = "Political Participation", y = "Predicted Probability of Gender (Female)",
       color = "Age Level") +
  theme_minimal()

```