---
title: "Midterm - Study Guide"
output: pdf_document
date: "2024-10-27"
---
# Bayesian (PS3)

## - Taking draws from prior / posterior
- Take draws from a distribution described by the parameters you've chosen as the prior
```{r}
set.seed(42)

# Beta
alpha_prior <- 1
beta_prior <- 1

beta_p_i <- rbeta(1048,alpha_prior , beta_prior)
beta_post_i <- rbeta(1048, alpha_prior + 558, beta_prior + 490)

# Gamma (PS3)
gamma_post_i <- rgamma(10000, shape = n + alpha, rate = beta + sum_log_data)

# Weibull (PS3)
data <- rweibull(n = 1000, shape = 5, scale = 7)

# Plot the prior
ggplot() + 
  geom_density(aes(x=gamma_post_i)) + 
  theme_minimal() +
  labs(title = "Posterior Distribution", x = "Theta", y = "Density")

```

## - Credible Interval
```{r}
# Calculate 95% credible interval
credible_interval <- quantile(gamma_post_i, probs = c(0.025, 0.975))
print(paste("95% Credible Interval: [", credible_interval[1], ", ", credible_interval[2], "]", sep = ""))
```

## - HPD interval
```{r}
# Calculate 95% HPD interval
hpd_interval <- bayestestR::hdi(gamma_post_i, ci = 0.95)
print(paste("95% HPD Interval: [", hpd_interval$CI_low, ", ", hpd_interval$CI_high, "]", sep = ""))
```


```{r}
# Gamma prior (Practice Midterm Bayesian problem)

x <- c(2, 2, 1, 3, 2, 1, 0, 4)
n <- length(x)

gamma_post_i <- rgamma(10000, shape = n + sum(x), rate = beta + n)

# Calculate 95% credible interval
credible_interval <- quantile(gamma_post_i, probs = c(0.025, 0.975))
print(paste("95% Credible Interval: [", credible_interval[1], ", ", credible_interval[2], "]", sep = ""))

# Calculate 95% HPD interval
hpd_interval <- bayestestR::hdi(gamma_post_i, ci = 0.95)
print(paste("95% HPD Interval: [", hpd_interval$CI_low, ", ", hpd_interval$CI_high, "]", sep = ""))
```

# Frequentist

## - Calcuate the 95% CI (CS12)

```{r , echo=FALSE}
x = c(1, 2, 4, 2, 3, 6, 2, 5)
y = c(3, 1, 7, 4, 5, 2, 3, 4)

z_value <- 1.96

x_mean <- mean(x)

se <- sd(x) / sqrt(length(x)-1)

lower_bound <- x_mean - z_value * se
upper_bound <- x_mean + z_value * se

ci <- c(lower_bound, upper_bound)
```

## - Calculate T & Z Critical Values

```{r}
# Define the parameters
alpha <- 0.05
n <- 22
df <- n - 1

# Calculate the critical value for a two-sided test
critical_value <- qt(1 - alpha / 2, df)

critical_value

# Calculate the critical value for a two-sided z-test
z_critical_value <- qnorm(1 - alpha / 2)

# Print the result
z_critical_value

```

# Non-Parametric

## - Bootstrap
```{r}
set.seed(42)

x <- c(2, 4, 5, 3, 8, 10, -2, 1, -1, 2, 5, 5)

replications <- 2000

bootstrap_medians <- rep(NA, replications)
for (i in 1:replications) {
  sample_x <- sample(x, length(x), replace = TRUE)
  bootstrap_medians[i] <- median(sample_x)
}

# Calculate Bootstrap SE
bootstrap_se <- sd(bootstrap_medians)

cat("Bootstrap SE for the median:", bootstrap_se, "\n")
```

```{r}
set.seed(42)

x <- c(2, 2, 1, 3, 2, 1, 0, 4)
replications <- 10000

bootstrap_means <- rep(NA, replications)
for (i in 1:replications) {
  sample_x <- sample(x, length(x), replace = TRUE)
  bootstrap_means[i] <- mean(sample_x)
}

# Calculate Bootstrap SE
bootstrap_se <- sd(bootstrap_means)

cat("Bootstrap SE for the mean:", bootstrap_se, "\n")

```

## - Jackknife
```{r}
n <- length(x)
jackknife_medians <- rep(NA, n)

for (i in 1:n) {
  jackknife_sample <- x[-i]
  jackknife_medians[i] <- median(jackknife_sample)
}

# Calculate the Jackknife estimate of the SE
jackknife_mean <- mean(jackknife_medians)
jackknife_se <- sqrt((n - 1) * mean((jackknife_medians - jackknife_mean)^2))

cat("Jackknife SE for the median:", jackknife_se, "\n")
```

# Null Hypothesis Testing

## - Calculate p-value
```{r}

x <- rnorm(60, mean=3, sd=1)
x.bar <- mean(x)
sd <- sd(x)

# Assuming known standard deviation from sample (σ)
n <- 60
mu_0 <- 0
sigma <- sd

pnorm(x.bar, mean = 0, sd = sd/sqrt(n), lower.tail=FALSE)
#print(1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n)))
```

### - Assume NH = True
```{r}
n = 40
mu = 0
null_mu = 0
sigma = 2

data = rnorm(n, mean = mu, sd = sigma)

print(1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n)))
```

### - Assume NH = True w/ sampling
```{r}
N = 400
n = 40
mu = 1
sigma = 2
null_mu = 0

ps = rep(0, N)
for (i in 1:N){
data = rnorm(n, mean = mu, sd = sigma)
ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
}
plot(density(ps))
```

### - Assume NH = False w/ sampling
```{r}
set.seed(1)

N = 400
n = 40
mu = 1
sigma = 2
null_mu = 0

ps = rep(0, N)
for (i in 1:N){
  data = rnorm(n, mean = mu, sd = sigma)
  ps[i] = 1-pnorm(mean(data),mean=null_mu,sd=sigma/sqrt(n))
}

print(sum(ps<=0.05)/N)

# Reject 94% of the time. This is called power.

plot(density(ps))
```

## - Likelihood Ratio Tests (CS15)

```{r}
# Likelihood under H0: mu = 0
L_H0 <- exp(sum(dnorm(x, mean = 0, sd = sigma, log = TRUE)))

# Likelihood under H1: mu = 20
L_H1 <- exp(sum(dnorm(x, mean = 20, sd = sigma, log = TRUE)))

# Calculate the Likelihood Ratio
lrt <- L_H0 / L_H1
lrt
```

# Calculating Power
```{r}
# Parameters
n <- 20  # number of insomniacs
p_null <- 0.8  # null hypothesis proportion
alpha_level <- 12  # rejection region threshold

# (a) Type I Error (α)
# Calculate the probability of Type I error
alpha <- pbinom(alpha_level, n, p_null)
cat("Type I Error (α):", alpha, "\n")

# (b) Type II Error (β) for p = 0.6
p_alternative_1 <- 0.6
beta_1 <- 1 - pbinom(alpha_level, n, p_alternative_1)
cat("Type II Error (β) when p = 0.6:", beta_1, "\n")

# (c) Type II Error (β) for p = 0.4
p_alternative_2 <- 0.4
beta_2 <- 1 - pbinom(alpha_level, n, p_alternative_2)
cat("Type II Error (β) when p = 0.4:", beta_2, "\n")

# (d) Power of the Test for each p value (loop)
p_values <- c(0.4, 0.5, 0.6, 0.7)
power_values <- numeric(length(p_values))

for (i in 1:length(p_values)) {
  beta <- 1 - pbinom(alpha_level, n, p_values[i])
  power_values[i] <- 1 - beta
}

# Display the results
for (i in 1:length(p_values)) {
  cat("Power of the Test when p =", p_values[i], ":", power_values[i], "\n")
}

# single call
p_value <- 0.4
beta <- 1 - pbinom(alpha_level, n, p_value)
power_value <- 1 - beta

# Optional: Plotting Power vs. p values
plot(p_values, power_values, type = "b", pch = 19, col = "blue",
     xlab = "True Proportion (p)", ylab = "Power",
     main = "Power of the Test vs. True Proportion")


```