---
title: "Midterm Revisions - DanCher"
author: "Dan Cher"
date: "2024-12-5"
output: pdf_document
---

# Problem 2f
Confidence Interval = [0.76, 1.095]

If we were to repeat this experiment many times and get a sample estimate each time, then 95% of the time our test-statistic will be between 0.76 and 1.095.

# Problem 4

```{r}
# Calculate power through simulations.

set.seed(42)

# Params
n <- 30
alpha <- 0.05
treatment_effect <- 0.4
simulations <- 1000
set.seed(42)

trial_sim <- function(n, treatment_effect) {
  control <- rnorm(n, mean = 0, sd = 1)
  treatment <- rnorm(n, mean = treatment_effect, sd = 1)
  
  test_result <- t.test(control, treatment, var.equal = TRUE)
  
  rejection_rule <- test_result$p.value < alpha
  
  return(rejection_rule)
}

rejections <- numeric(simulations)
for (i in 1:simulations) {
  rejections[i] <- trial_sim(n, treatment_effect)
}

power <- mean(rejections)
cat("Power :", power, "\n")

```

```{r}
sd <- 1           # Standard deviation
effect_size <- 0.4  # Standardized effect size
n <- 30            # Sample size

# Calculate standard error (SE)


# Critical z-value for two-tailed test at alpha = 0.05
z_critical <- 1.96

power <- (1 - pnorm(z_critical - effect_size / SE)) + 
  pnorm(-z_critical - effect_size / SE)

# Power calculation (simplified)
SE <- sd / sqrt(n)
power <- 1 - pnorm(1.96, mean = effect_size / SE, sd = 1) + pnorm(-1.96, mean = effect_size / SE, sd = 1)

cat(pnorm(1.96, mean = effect_size / SE, sd = 1), "\n")
cat(pnorm(-1.96, mean = effect_size / SE, sd = 1))

```

```{r}
# Parameters
n <- 30           # Sample size
alpha <- 0.05     # Significance level
effect_size <- 0.4  # Standardized effect size
sd <- 1            # Standard deviation

power <- pnorm(1.96 * sqrt(2 / n), mean = 0.4, sd = sqrt(2 / n))

SE <- sd / sqrt(n)
power <- 1 - pnorm(1.96, mean = sd / sqrt(n), sd = 1) + pnorm(-1.96, mean = effect_size / SE, sd = 1)
```

```{r}
# Power Analysis for Treatment Effect

# Set parameters
effect_size <- 0.4  # Standardized effect size
n <- 30             # Sample size
alpha <- 0.05       # Significance level

# Calculate critical value for two-sided test
critical_value <- qt(1 - alpha/2, df = 2*n - 2)

# Calculate non-centrality parameter
ncp <- sqrt(n/2) * effect_size

# Calculate power
power <- 1 - pt(critical_value, df = 2*n - 2, ncp = ncp) + 
         pt(-critical_value, df = 2*n - 2, ncp = ncp)

# Print results
cat("Power Analysis Results:\n")
cat("Effect Size:", effect_size, "\n")
cat("Sample Size:", n, "\n")
cat("Significance Level (α):", alpha, "\n")
cat("Power:", power, "\n")
```


# Problem 5

```{r}

typeI_sim <- function(n) {
  control <- rnorm(n, mean = 0, sd = 1)
  treatment <- rnorm(n, mean = 0, sd = 1)
  
  test_result <- t.test(control, treatment, var.equal = TRUE)
  
  rejection_rule <- test_result$p.value < alpha
  
  return(rejection_rule)
}

rejections <- numeric(simulations)
for (i in 1:simulations) {
  rejections[i] <- typeI_sim(n)
}

type_I_error_rate <- mean(rejections)

type_II_error_rate <- 1 - power

# Output the results
cat("Type I Error Rate:", type_I_error_rate, "\n")
cat("Type II Error Rate:", type_II_error_rate, "\n")
```

Type I error rate is the probability of rejecting the null hypothesis when it is true. In this instance that would be the probability of concluding there is a significant treatment effect when there actually isn't. Given our inputs, we see that 5.5% of the time, we incorrectly conclude there is a treatment effect when there actually isn't one. 

Type II error rate is the probability of failing to reject the null hypothesis when it is false. We fail to detect a significant effect (effect = 0.4 in our case) when it does actually exist. Given our inputs, 68% of the simulations show the test failed to reject the null hypothesis despite there being a true effect.

